\documentclass[a4paper, notitlepage, 10pt]{article}
\usepackage{geometry}
\usepackage{titling}
\usepackage{authblk}

% Page Layout Configuration
\geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
}

% Font and Typography
\usepackage{setspace,relsize}
\usepackage{moreverb}

% Tables, Figures, and Listings
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usepackage{wrapfig}
\usepackage{listings}

% Mathematical Packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bbm}

% Text Formatting
\usepackage{indentfirst}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{cancel}
\usepackage[normalem]{ulem}

% Hyperlinks and Color
\definecolor{keywords}{RGB}{255,0,90}
\definecolor{comments}{RGB}{0,0,113}
\definecolor{red}{RGB}{160,0,0}
\definecolor{green}{RGB}{0,150,0}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,      % Internal links
    citecolor=blue,      % Citations
    urlcolor=blue        % External hyperlinks
}
\usepackage[square,numbers]{natbib}

% Code Listings Style
\lstset{
    language=Python, 
    basicstyle=\ttfamily\footnotesize, 
    keywordstyle=\color{blue},
    commentstyle=\color{green},
    stringstyle=\color{red},
    showstringspaces=false
}

% Title and Authors
\title{\Large \bfseries Mean Field Optimal Control Applied to McKean-Vlasov and Fokker-Planck Models}
\date{\today}
\author{Lucas Machado Moschen}
\affil{\small Department of Mathematics, Imperial College London}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% Mathematical Definitions %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\R}{\mathbb{R}}
\newcommand{\bb}[1]{\boldsymbol{#1}}

\newcommand{\ev}{\mathbb{E}}
\newcommand{\var}{\operatorname{Var}}
\newcommand{\cor}{\operatorname{Cor}}
\newcommand{\cov}{\operatorname{Cov}}
\newcommand{\almost}{\text{a.e. }}

\newcommand{\ind}{\mathbbm{1}}

% Theorem Environments

\newtheorem{theorem}{Theorem}[section]
\newtheorem{claim}{Claim}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{lemma}{Lemma}[section]

\theoremstyle{definition}
\newtheorem{example}{Example}[section]

\newtheoremstyle{customstyle} % Define a custom style for assumptions
  {10pt}      % Space above
  {10pt}      % Space below
  {\normalfont\setlength{\parindent}{0pt}\setlength{\leftskip}{2em}}
  {}          % Indent amount
  {\bfseries} % Theorem head font
  {}          % Punctuation after theorem head
  { }         % Space after theorem head
  {\thmnumber{#2}}

\theoremstyle{customstyle}
\newtheorem{assumption}{}
\renewcommand{\theassumption}{(A\arabic{assumption})} 

\providecommand{\assumptionautorefname}{Assumption}
\newcommand{\lemmaautorefname}{Lemma}
\newcommand{\remarkautorefname}{Remark}
\newcommand{\corollaryautorefname}{Corollary}
\newcommand{\definitionautorefname}{Definition}
\newcommand{\propositionautorefname}{Proposition}
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Subsection}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\newcommand{\LM}[1]{\textcolor{red}{#1}}
\renewcommand{\citeauthoryear}[1]{\citeauthor{#1} (\citeyear{#1})}

\begin{document}

\maketitle

%\tableofcontents

\section{Introduction}

\subsection{Mean-Field Optimal Control}

Mean-field Optimal Control emerged from the work of \citeauthoryear{Fornasier2013}. 
Its primary goal is to model and control systems with many interacting agents. 
Mean-field models, which consider that each individual is influenced by the overall average behaviour of the entire population or the closest neighbours, describe the collective dynamics of such systems. 
This concept draws from models of particle interaction in physics, that use binary forces to mimic the system behaviour.

In contrast to Mean-field Games, where agents operate independently, in mean-field optimal control the agents are governed by a {\em centralized governor} or {\em controller}. 
This approach has several applications across diverse fields: in economics, it can model market behaviour; in physics, it describes the dynamics of particle systems; in biology, it can capture collective animal behaviour; and in social science, it can represent opinion dynamics.
Mean-field optimal control extends classical control theory to incorporate population-level interactions, resulting in optimal control problems constrained by partial differential equations (PDEs).

When introduced by \citeauthoryear{Fornasier2013}, the authors demonstrated rigorously that the limit of a system of multiple agents, modelled by a finite-dimensional optimal control problem with ordinary differential equations (ODEs), leads to an infinite-dimensional optimal control problem constrained by a Vlasov-type PDE~\cite{Fornasier2013}. 
The model governing the agents is expressed as
\begin{equation}\label{eq:ode_model}
    \begin{cases}
        \dot{x}_i &= v_i \\
        \dot{v}_i &= (H * \mu_N)(x_i, v_i) + u(t,x_i,v_i), \quad i=1,\dots,N, \quad t \in [0,T],
    \end{cases}
\end{equation}
where 
\[
\mu_N = \frac{1}{N} \sum_{j=1}^N \delta_{(x_i,v_i)}
\]
is the empirical measure of the positions and velocities of the $N$ agents in $\R^{2d}$.
The function $u$ represents the control exerted over all agents from a central authority, which aims to minimize the cost functional
\[
J^{N}(u) := \int_0^T \int_{\R^{2d}} (L(x,v,\mu_N) + \psi(u(t,x,v)) \, d\mu_N(t,x,v) \, dt,
\]
where $\psi: \R^d \to [0, \infty)$ is a convex function, and $u$ is a Carathéodory function with sublinear growth in the variables $x$ and $v$.

This review focuses on the following topics in the literature:

\begin{itemize}
    \item Existence and uniqueness results for McKean-Vlasov and Fokker-Planck models.
    \item The well-posedness of these models.
    \item Numerical methods for mean-field control problems.
    \item Extensions of Pontryagin's Maximum Principle to mean-field settings.
    \item The integration of machine learning techniques in mean-field optimal control.
\end{itemize}

Another interesting feature that we might want to go further is the relation between Mean-field Games and control of McKean-Vlasov equations.
The ideas come from~\citeauthoryear{carmona2013control} who compared both methods.
In particular, they explain the difference between the nature and solutions to the two approaches by investigating the corresponding forward-backward systems.

\section{Mathematical Background}

\subsection{Convergence}

In~\cite{Fornasier2013}, they theoretically prove that~\eqref{eq:ode_model} converges for $N \to \infty$ to a solution of the PDE constrained problem
\[
\frac{\partial \mu}{\partial t} + v \cdot \nabla_x \mu = \nabla_v \cdot [(H * \mu + u)\mu],
\]
with cost functional 
\[
J(u) := \int_0^T \int_{\R^{2d}} (L(x,v,\mu) + \psi(u(t,x,v))) \, d\mu(t,x,v) \, dt,
\]
where $\mu$ is a probability measure valued weak solution to the PDE.

\subsection{The mean-field models}

By the McKean-Vlasov equation, we may also mean the stochastic differential equation (SDE)
\[
dX_t = b(t, X_t, \mathbb{P}_{X_t}) dt + \sigma(t, X_t, \mathbb{P}_{X_t}) dW_t,
\]
which describes the dynamics of an average particle under a drift term $b$, a diffusion term $\sigma$, a Brownian motion $W_t$, and a population probability distribution $\mathbb{P}_{X_t}$.

Alternatively, the Fokker-Planck equation describes the evolution of the probability density function $\rho(t, x)$ of a stochastic process $X_t$:
\[
\frac{\partial \rho}{\partial t} = -\nabla_x \cdot [b(t, x, \rho) \rho] + \frac{1}{2} \nabla_x^2 : [\sigma \sigma^\top (t, x, \rho) \rho].
\]

This PDE is the macroscopic counterpart to the microscopic dynamics described by the SDE.

\subsection{Stochastic interactions}

By extending model~\eqref{eq:ode_model} to include random fluctuations, we consider the model. 
\[
dX_t^i = - \nabla H^N(X_t^1, \dots, H_t^N) \, dt + \sqrt{2\sigma} \, dB_t^i. 
\]
The distribution of the $N$-particle positions satisfies a Fokker-Planck equation given by
\[
\partial_t \rho^N_t = \nabla_N \cdot \left(\nabla H^N \rho_t^N \right) + \sigma \Delta \rho_t^N,
\]
where $\sigma$ is the system's temperature.
By assuming the exchangeability of the particles, we get the McKean-Vlasov PDE.
\[
\partial_t \rho = \nabla \cdot \left(\rho[(\nabla W * \rho) + \nabla V]\right) + \sigma \Delta \rho
\]

\section{Literature Review}

\subsection{Existence and Uniqueness Results}

\subsubsection{Analytical Techniques}

The existence and uniqueness of solutions to McKean-Vlasov equations and associated control problems are established. 

\subsection{Well-posedness Analysis}

\subsubsection{Theoretical Developments}

\subsubsection{Machine Learning Integration}

\begin{itemize}
    \item \textbf{Neural Lyapunov Functions}: The paper from \citeauthoryear{gaby2022lyapunov} introduces a deep neural network, called {\em Lyapunov-Net}, to approximate Lyapunov functions for high-dimensional systems. 
    These functions can then be used to check the stability of systems and must meet two conditions: they must be positive definite and have a negative orbital derivative~\cite{gaby2022lyapunov}.
    The main goal of their work is to make a method that works better and faster than older methods, especially when dealing with systems that have many dimensions. 

    A Lyapunov-Net is defined using two functions, where $V_\theta(x) = |\phi_\theta(x) - \phi_\theta(0)| + \bar{\alpha} \|x\|$. 
    This setup ensures $V_\theta$ is always positive definite. The network is trained to minimize a risk function:
    \[
    \ell(\theta) = \int_{\Omega} (DV_{\theta}(x) \cdot f(x) + \gamma \|x\|)_+^2 \, dx,
    \]
    where $\gamma$ is a chosen parameter and $\Omega$ represents the space where the problem exists. 
    The function focuses on making sure that the $D V_{\theta}$ along the system trajectories is negative, which is required for stability. 
    For this integral, Monte Carlo integration was performed.
    
    The paper concludes that Lyapunov-Net is better at approximating Lyapunov functions than current methods and does this more efficiently. 
    However, there are some issues near points of equilibrium due to the risk function used. 
    Future research may look into better network designs, more detailed theoretical support, and applying the method to more complex systems. 
    This could help us learn more widely in system stability and control fields.

    \item \textbf{Stability Analysis via Data-Driven Methods}: Machine learning models analyze the stability of equilibria using simulation data.
\end{itemize}


\subsection{Numerical Methods for Mean Field Optimal Control}

\subsubsection{Classical Numerical Methods}

Traditional numerical methods for MFOC involve discretization techniques for PDEs and SDEs, such as {\em finite difference methods}, {\em finite element methods}, {\em spectral methods}, and {\em Monte Carlo simulations} of paths.

\subsubsection{Machine Learning Approaches}

Recent research has explored the use of machine learning to overcome the limitations of classical methods: Physics-informed neural networks (PINN), Reinforcement Learning, Generative Models for Distribution Approximation, etc.

\subsection{Optimality conditions}

\subsubsection{Classical Applications}

PMP has been extended to mean field models, leading to necessary conditions that involve both the state and the distribution~\cite{aronna2021}

\subsubsection{Machine Learning Enhancements}

\begin{itemize}
    \item \textbf{Neural Network Approximation of Adjoint States}: Neural networks approximate the adjoint processes in the stochastic PMP.
\end{itemize}

\subsection{Open questions}

\begin{itemize}
    \item \textbf{Curse of Dimensionality}: Classical numerical methods struggle with high-dimensional state spaces.
    \item \textbf{Machine Learning and control}: Deep learning models can handle high-dimensional inputs, mitigating the curse of dimensionality.
    \item Understanding the basis of attractions of the McKean-Vlasov equation.
\end{itemize}

\section{Conclusion}

\subsection{Summary of key points}

\subsection{Gaps in the literature}

\subsection{Suggestions for future research from the papers}

\subsection{Final remarks}

\newpage

\section{Problem settings}

Consider a system with \( N \) exchangeable agents, each located at position \( x_t^i \) (for \( i = 1, \dots, N \)) at time \( t \), evolving according to the following equation:
\[
dx_t^i = - \left( \nabla V(x_t^i) + \frac{1}{N} \sum_{j=1}^N \nabla W(x_t^i - x_t^j) \right) dt + \sqrt{2 \sigma} \, dB_t^i,
\]
where \( V \) represents a confining potential that influences each agent’s position, \( W \) is a communication potential that models the interaction between agents, the diffusion coefficient \( \sigma \) denotes the system's temperature, and \( \{B_t^i\}_{t \ge 0} \) are independent standard Brownian motions associated with each agent, introducing stochastic fluctuations in the system.
This system evolves within a torus, ensuring periodic boundary conditions on agent positions.

Under some regularity conditions, the marginal distribution of the position of a single agent converges as \( N \to \infty \) to a probability density \( \mu \) that satisfies the McKean–Vlasov equation
\[
\partial_t \mu = \nabla \cdot \left( \mu[(\nabla W * \mu) + \nabla V] \right) + \sigma \Delta \mu,
\]
where \( \mu \) is defined on the torus $\Omega$ with periodic boundary conditions.
The initial distribution of \( \mu \) is given by \( \mu(\cdot, 0) = \mu_0 \). 
Here, the interaction term \( (W * \mu)(x) \) is defined as
\[
(W * \mu)(x) := \int_{\Omega} W(x - y) \, \mu(y) \, dy
\]
over the torus, representing the convolution of the communication function \( W \) with the density \( \mu \).

As studied by \citeauthoryear{Carrillo2020}, the McKean–Vlasov equation on the torus exhibits global asymptotic stability of solutions towards a homogeneous steady state. 
Additionally, the equation allows for nontrivial solutions that can exhibit an infinite number of bifurcations, depending on the interaction potential. 
The authors establish sufficient conditions under which both continuous and discontinuous phase transitions can occur~\cite{Carrillo2020}.
In their analysis, the authors apply these theoretical results to a range of interaction potentials used in models like the noisy Kuramoto, Keller–Segel, and noisy Hegselmann-Krausse models, exploring their implications for synchronization, chemotaxis, and opinion dynamics respectively.

Suppose that $W$ is convex, even, and with polynomial growth and that $V$ is uniformly convex, i.e., for some $\beta > 0$, $\nabla^2 V(x) - \beta I$ is a positive-semidefinite matrix.
Under these conditions, \citeauthoryear{malrieu2001logarithmic} demonstrates that there exist a unique function $\bar{\mu}$ and a constant $K > 0$ such that for all $t >0$, $\|\mu(t,\cdot) - \bar{\mu}\|_{L^1} \le Ke^{-\beta t/2}$, where $\bar{\mu}$ is the unique stationary solution satisfying the detailed balance condition
\[
\bar{\mu}(x) = \frac{1}{Z} e^{-\frac{1}{\sigma}\left(V(x) - W * \bar{\mu}(x)\right)}
\]
with $Z = \int_{\Omega} e^{-\frac{1}{\sigma}\left(V(y) - W * \bar{\mu}(y)\right)} \, dy$.

We are interested in {\bf controlling} the dynamics described by this equation to accelerate the convergence of $\mu(t,\cdot)$ to the equilibrium of the system, especially in the cases where $\beta$ is small.
We then take 
\[
V(x,t) = G(x) + \sum_{j=1}^{m} \alpha_j(x) u_j(t)
\]
for uniformly convex functions $G$ and $m$ functions $\alpha_j$.
The functions $\alpha_j$ act as a control shape function determining the regions in the space where the time-dependent control function $u_j$ will influence.
We then introduce the controlled dynamics 
\begin{equation}
    \label{eq:controlled_mckean_vlasov}
    \partial_t \mu = \nabla \cdot \left( \mu \left[(\nabla W * \mu) + \nabla G + \sum_{j=1}^m u_j \nabla \alpha_j \right] \right) + \sigma \Delta \mu.
\end{equation}
This formulation is similar to~\cite{Breiten2018}, in which they consider the Fokker-Planck equation as the state dynamics.
\LM{The function $u$ needs to be chosen to guarantee that $\bar{\mu}$ still is the unique equilibrium of the system}.

\LM{XXX: discuss the well-posedness of the equation, (I) existence and uniqueness of well-behaved solutions; (II) properties that $\mu$ is a PDF}.
\LM{What guarantees that for any $u$, there will be only one equilibrium? Is it enough to have $u \to 0$? In which sense?}

The abstract formulation of equation~\eqref{eq:controlled_mckean_vlasov} using operators is
\[
\dot{\mu} = \mathcal{A} \mu + \sum_{j=1}^m u_j(t) \mathcal{N}_j \mu +  \mathcal{W}(\mu),
\]
where $\mathcal{A} \mu = \nabla \cdot (\mu \nabla G + \sigma \nabla \mu)$, $\mathcal{N}_j \mu = \nabla \cdot (\mu \nabla \alpha_j)$, and $\mathcal{W}(\mu) = \nabla \cdot \left( \mu \left( \nabla W * \mu \right) \right)$ captures the nonlinear and non-local interaction term involving the convolution \(\nabla W * \mu\).

Define $y = \mu - \bar\mu$ so that we have
\begin{equation}
    \label{eq:abstract_formulation_mckean_vlasov}
    \dot{y} = \mathcal{A} y + \mathcal{W}(y+\bar{\mu}) - \mathcal{W}(\bar{\mu}) + \sum_{j=1}^m u_j(t) \mathcal{N}_j y + \mathcal{B}_j u_j(t),
\end{equation}
where $\mathcal{B}_j = \mathcal{N}_j \bar\mu$.

To compute the linearised version of equation~\eqref{eq:abstract_formulation_mckean_vlasov}, the first step is to linearized $\mathcal{W}$ around $\bar{\mu}$ through its Fréchet derivative. 
Notice that 
\[
\begin{split}
    \mathcal{W}(y+\bar{\mu}) - \mathcal{W}(\bar{\mu}) &= \nabla \cdot \left[(y+\bar\mu)(\nabla W * y + \nabla W * \bar\mu) - \bar\mu (\nabla W * \bar{\mu})\right] \\
    &= \nabla \cdot [(y+\bar\mu)(\nabla W * y) + y(\nabla W * \bar{\mu})] \\
    &= \nabla \cdot [\bar\mu(\nabla W * y) +  y(\nabla W * \bar{\mu})] + \nabla \cdot [y(\nabla W * y)] \\
    &\approx \nabla \cdot [\bar\mu(\nabla W * y) +  y(\nabla W * \bar{\mu})] \\ 
    &=: \mathcal{D}_W y,
\end{split}
\]
where $\mathcal{D}_W y$ is the functional derivative of $\mathcal{W}$.
The linearized system is then
\begin{equation}
    \label{eq:linearized_abstract_mckean_vlasov}
    \dot{y} = (\mathcal{A} + \mathcal{D}_W + \delta I)y + \sum_{j=1}^m \mathcal{B}_j u_j(t),
\end{equation}
where we include a parameter $\delta > 0$ to guarantee that all eigenvalues are negative, and $\mathcal{B} = (\mathcal{B}_1, \dots, \mathcal{B}_m)$ and $u(t) = (u_1(t), \dots, u_m(t))$. 
We formulate this as an optimization problem where the objective is to minimize the integral of the squared \(L^2\) norm of \(y\) over time, together with the squared magnitude of the control function \(u(t)\), explicitly stated as
\begin{equation}
    \label{eq:quadratic_cost_functional}
    \min_u \frac{1}{2}\int_0^{\infty} \left( \langle y(t), \mathcal{M} y(t) \rangle_{L^2(\Omega)} + \|u(t)\|_{2}^2\right) dt,
\end{equation}
where $\mathcal{M}$ is a self-adjoint nonnegative operator.

The cost functional written in equation~\eqref{eq:quadratic_cost_functional} is associated to the {\em Ricatti equation} in the weak sense
\[
\langle (\mathcal{A} + \mathcal{D}_W + \delta I)^* \Pi y_1, y_2 \rangle_{L^2(\Omega)} + \langle \Pi (\mathcal{A} + \mathcal{D}_W + \delta I)y_1, y_2 \rangle_{L^2(\Omega)} - \langle \mathcal{B}^* \Pi y_1, \mathcal{B}^* \Pi y_2 \rangle_{L^2(\Omega)} + \langle \mathcal{M} y_1, y_2 \rangle_{L^2(\Omega)}= 0,
\]
for all functions $y_1, y_2$ in the domain of $\mathcal{A}$.
If $(\mathcal{A} + \mathcal{D}_W, \mathcal{B})$ is $\delta$-stabilizable (\LM{XXX: this means that there exists a linear feedback control such that the resulting system will have with eigenvalues with real parts being at most $-\delta$.}), there exists a unique self-adjoint $\Pi$ such that $\mathcal{A} + \mathcal{D}_W + \delta I - \mathcal{B} \mathcal{B}^* \Pi$ generates an exponentially stable semigroup.
Additionally, it suggests that we need a particular choice of $\alpha$ so that the pair is $\delta$-stabilizable for $\delta \ge -\rho(\mathcal{A}^*)$.
\LM{XXX: we need to find such a condition for our case, I think it will be less trivial}.
We then apply to equation~\eqref{eq:abstract_formulation_mckean_vlasov} the feedback law $u = -\mathcal{B}^* \Pi y$. \LM{XXX: we should then study the local stability here.}

\subsection{Numerical Approach: Spectral Method and Riccati equation}

To numerically solve the controlled McKean–Vlasov equation~\eqref{eq:controlled_mckean_vlasov}, we employ a spectral method due to its accuracy and efficiency for periodic boundary conditions on the torus. 
Given that the domain is periodic, Fourier series are a natural choice for the basis functions. 
To obtain the control function, after discretizing the linearized system using the spectral method, we will have a finite-dimensional linear-quadratic control problem which will provide, through a Riccati equation, an optimal feedback control formula.
We then will plug this control into the original non-linear equation and solve the system of ODEs, which will lead to the approximated controlled McKean–Vlasov equation. 
\LM{We should obtain some approximation results such as in \url{https://core.ac.uk/download/pdf/82197863.pdf}. 
For instance, they demonstrated that the approximated feedback control matrix converges to the abstract operator.}

We approximate \( y(x,t) \) and \( \bar{\mu}(x) \) using truncated Fourier series expansions:
\[
y(x,t) \approx \hat{y}_L(x,t) := \sum_{k=-L}^{L} a_k(t) \phi_k(x), \quad \bar{\mu}(x) \approx \hat{\mu}_L(x) := \sum_{k=-L}^{L} \bar{\mu}_k \phi_k(x),
\]
where \(\phi_k(x)\) are the Fourier modes, \( k \in \mathbb{Z} \), and \( L \) is a truncation parameter that determines the number of modes used in the approximation.
For instance, in one dimension, when the torus is $[0,2\pi)$, we can consider the functions $\phi_k(x) = e^{ikx}$ with $a_{-k} = \overline{a_k}$, or purely sines and cosines.
Our goal is to project the controlled McKean–Vlasov equation~\eqref{eq:controlled_mckean_vlasov} onto the finite-dimensional subspace spanned by the chosen Fourier modes, resulting in a system of $2L+1$ ordinary differential equations (ODEs) for the coefficients \( a_k(t) \).

First, we compute $\hat\mu_L$ and project the initial condition on the Fourier basis.
We approximate the equation that $\bar\mu$ satisfies, leading to
\[
\sum_{k=-L}^{L} \bar\mu_k \phi_k(x) = \frac{1}{\hat{Z}} e^{-\sigma^{-1}G(x) - \sigma^{-1}\sum_k \bar\mu_k W * \phi_k(x)},
\]
for some constant $\hat{Z}$ that normalizes $\bar\mu$ to a probability density.
Integrating against $\phi_i^*$ we obtain,
\[
\bar\mu_i = \frac{1}{\hat{Z}\| \phi_i \|^2} \int_{\Omega} e^{-\sigma^{-1}G(x) - \sigma^{-1}\sum_{k=-L}^L \bar\mu_k W * \phi_k(x)} \phi_i^*(x) \, dx,
\]
which results in a set of $2L+1$ non-linear equations.
For the initial condition, set
\[
\mu_0(x) \approx \hat\mu_0(x) := \sum_{k=-L}^L \bar \mu_{k,0} \phi_k(x) \implies \bar\mu_{i,0} = \frac{1}{\|\phi_i\|^2}\int_{\Omega} \mu_0(x) \phi_i^*(x) \, dx. 
\]

Starting with the discretisation of the linearized McKean–Vlasov equation as outlined in equation~\eqref{eq:linearized_abstract_mckean_vlasov}, we utilize the Fourier basis functions as test functions. 
This yields, for each $-L \le i \le L$, the discretized version of the linearized equation:
\[
\begin{split}
    \sum_{k=-L}^{L} \dot{a}_k(t) \int_{\Omega} \phi_k(x) \phi_i^*(x) \, dx &= -\sum_{k=-L}^{L} a_k(t) \left( \int_{\Omega} \phi_k(x) \nabla G(x) \cdot \nabla \phi_i^*(x)  + \sigma \nabla \phi_k(x) \cdot \nabla \phi_i^*(x) \, dx \right) \\
    - \sum_{k=-L}^{L} &a_k(t) \left( \int_{\Omega} \hat\mu_L(x)( \nabla W * \phi_k(x) ) \cdot \nabla \phi_i^*(x) + \phi_k(x) (\nabla W * \hat\mu_L(x)) \cdot \nabla \phi_i^*(x) \, dx \right) \\ 
    +\delta \sum_{k=-L}^{L} &a_k(t) \int_{\Omega} \phi_k(x) \phi_i^*(x) \, dx - \sum_{j=1}^m u_j(t) \int_{\Omega} \hat\mu_L(x) \nabla \alpha_j(x) \cdot \nabla \phi_i^*(x) \, dx. 
\end{split}
\]

The above system can be written in matrix form as 
\[
\dot{a}(t) = -\Phi^{-1}(L_G + \sigma D + K - \delta \Phi) a(t) + \Phi^{-1}B u(t), 
\]
with $\Phi_{ik} = \langle \phi_k, \phi_i \rangle$, $[L_G]_{ik} = \langle \phi_k \nabla G, \nabla \phi_i \rangle$, $D_{ik} = \langle \nabla \phi_k, \nabla \phi_i \rangle$, $K_{ik} = \langle \hat\mu_L(\nabla W * \phi_k) + \phi_k(\nabla W * \hat\mu_L), \nabla \phi_i \rangle$, and $B_{ij} = -\langle \hat\mu_L\nabla \alpha_j, \nabla \phi_i \rangle$.
By choosing an orthonormal basis, we can set $\Phi = I$ and $A = -(L_G + \sigma D + K - \delta \Phi)$.
Moreover, the cost functional reads as 
\[
\frac{1}{2} \int_0^{\infty} a(t)^T M a(t) + \|u(t)\|_2^2 \, dt,
\]
where $M_{ij} = \langle \phi_i, \mathcal{M} \phi_j \rangle$.
The associated Riccati equation is, then, 
\[
A^* \hat\Pi + \hat\Pi A - \hat\Pi B B^* \hat\Pi + M = 0,
\]
which can be solved numerically, leading to the feedback control $u(t) = -B^* \Pi a(t)$.

We also perform the same operation for the non-linear equation~\eqref{eq:abstract_formulation_mckean_vlasov} with $u(t)$ in the feedback form.
This leads to 
\[
\begin{split}
    \sum_{k=-L}^{L} \dot{a}_k(t) \int_{\Omega} \phi_k(x) \phi_i^*(x) \, dx &= -\sum_{k=-L}^{L} a_k(t) \left( \int_{\Omega} \phi_k(x) \nabla G(x) \cdot \nabla \phi_i^*(x)  + \sigma \nabla \phi_k(x) \cdot \nabla \phi_i^*(x) \, dx \right) \\
    - \sum_{k=-L}^{L} &a_k(t) \left( \int_{\Omega} \hat\mu_L(x)( \nabla W * \phi_k(x) ) \cdot \nabla \phi_i^*(x) + \phi_k(x) (\nabla W * \hat\mu_L(x)) \cdot \nabla \phi_i^*(x) \, dx \right) \\
    - \sum_{k=-L}^{L} &\sum_{l=-L}^{L} a_k(t) a_l(t) \int_{\Omega} \phi_k(x)( \nabla W * \phi_l(x)) \cdot \nabla \phi_i^*(x) \, dx \\
    + \sum_{j=1}^m (B^* \Pi &a(t))_j \sum_{k=-L}^{L} a_k(t) \int_{\Omega} \phi_k(x) \nabla \alpha_j(x) \cdot \nabla \phi_i^*(x) \, dx \\
    + (B^* \Pi &a(t))_j \int_{\Omega} \hat\mu_L(x) \nabla \alpha_j(x) \cdot \nabla \phi_i^*(x) \, dx. 
\end{split}
\]
The above system can be rewritten as 
\[
\dot{a}(t) = -(L_G + \sigma D + K) a(t) - T : (a(t) \otimes a(t)) + \sum_{j=1}^m (B_j^* \Pi a(t)) \Psi^{(j)} a(t)  -  B B^* \Pi a(t),
\]
where $T_{ikl} = \langle \phi_k \nabla W * \phi_l, \nabla \phi_i \rangle$, $\Psi_{ik}^{(j)} = \langle \phi_k \nabla \alpha_j, \nabla \phi_i \rangle$, $B_j$ is the $j$-th column of $B$,  and $D:C$ is a double contraction between the tensors $D$ and $C$.
By solving this system of ODEs with $a_k(0) = \bar\mu_{i,0}$, we obtain an approximation for the controlled McKean-Vlasov equation.

\subsubsection{1-D case}

In one dimension, over the interval $[0,d]$ with periodic boundary conditions, consider the orthonormal basis
\[
\phi_k(x) = \frac{1}{\sqrt{d}} e^{\frac{2\pi}{d}i k x}.
\]

An easy way to project a function on the Fourier basis uses the so-called Fast Fourier Transform.
If over an interval $[0,d]$,
\[
g(x) = \sum_{k=-L}^L c_k \phi_k(x),
\]
then we know that
\[
c_k = \sqrt{\frac{1}{d}} \int_0^d f(x) e^{-\frac{2\pi}{d} k x i} \, dx \approx \frac{d}{N\sqrt{d}} \sum_{j=0}^{N-1} f(t_j) e^{-\frac{2\pi}{d} k t_j i}
\]
for points $0 \le t_0 < t_1 < \cdots < t_{N-1} \le d$. 
We consider the mid interval points $t_j = (j+1/2) d/N$, so
\[
\begin{split}
    c_k &\approx \frac{\sqrt{d}}{N} \sum_{j=0}^{N-1} f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)e^{-\frac{2\pi}{N} k (j+1/2) i} \\
    &= \frac{\sqrt{d}}{N} e^{-\frac{\pi}{N} k i}\sum_{j=0}^{N-1} f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)e^{-\frac{2\pi}{N} j k i} \\
    &=  \frac{\sqrt{d}}{N} e^{-\frac{\pi}{N} k i} FFT_k[(f_0, \dots, f_{N-1})],
\end{split}
\]
where $f_j = f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)$, for $k \ge 0$.
For $k < 0$, notice that $c_k = \overline{c_{-k}}$.

Let us calculate $W * \phi_k$ for a periodic function $W$.
Write 
\[
W(x) = \sum_{j=-\infty}^{\infty} w_j \phi_j (x).
\]
Then,
\[
W * \phi_k(x) = \sum_{j=-\infty}^{\infty} w_j \int_{\Omega} \phi_j(x-y) \phi_k(y) \, dy = \sum_{j=-\infty}^{\infty} w_j  \phi_j * \phi_k(x) = \sqrt{d} w_k \phi_k(x)
\]
where $w_k$ can be computed similarly using the FFT, as for $\mu_0$.

Using this fact to calculate $\bar\mu_i$,
\[
\bar\mu_i = \frac{1}{\hat{Z}} \int_{\Omega} \exp\left\{-\frac{G(x)}{\sigma} - \frac{\sqrt{d}}{\sigma}\sum_{k=-L}^L \bar\mu_k w_k \phi_k(x)\right\} \phi_i^*(x) \, dx.
\]

Let us find a better way to compute $L_G$.
Notice that \( \phi_k'(x) = \frac{2\pi}{d} ik \phi_k(x) \).
By definition,
\[
[L_G]_{kl} = \int_{\Omega} \phi_l(x) G'(x) \overline{\phi_k'(x)} \,dx = -\frac{2\pi}{d} i k \int_{\Omega} \phi_l(x) G'(x) \phi_{-k}(x) \,dx.
\]
Write 
\[
G(x) = \sum_{j=-\infty}^{\infty} g_j \phi_j(x) \implies G'(x) = \sum_{j=-\infty}^{\infty} g_j \phi_j'(x) = \sum_{j=-\infty}^{\infty} \frac{2\pi}{d} ij g_j \phi_j(x).
\]
Then, we get that 
\[
\begin{split}
    [L_G]_{kl} &= \frac{4\pi^2}{d^2} k \sum_{j=-\infty}^{\infty} j g_j \int_{\Omega} \phi_l(x) \phi_j(x) \phi_{-k}(x) \, dx \\
    &= \frac{4\pi^2}{d^2\sqrt{d}} k  \sum_{j=-\infty}^{\infty} j g_j \int_{\Omega} \frac{1}{\sqrt{d}} e^{\frac{2\pi}{d} i (l+j-k)x} \, dx \\
    &= \frac{4\pi^2}{d^2\sqrt{d}} k (k-l) g_{k-l}.
\end{split}
\]

We move to compute matrix $D$, which is given by
\[
D_{kl} = \int_{\Omega} \phi_l'(x) \overline{\phi_k'(x)} \, dx = \frac{4 \pi^2}{d^2} kl \int_{\Omega} \phi_{l}(x) \phi_{-k}(x) \, dx = \frac{4 \pi^2}{d^2} kl \mathbbm{1}_{k=l}.
\]

Let us do this for matrix $K$.
By definition, 
\[
\begin{split}
    K_{kl} &= \int_{\Omega} [\bar{\mu}_L(x) (W' * \phi_l(x)) + \phi_l(x) (W'*\bar\mu_L(x))] \overline{\phi_k'(x)} \, dx \\
    &= \sum_{j=-L}^{L} \bar\mu_j \int_{\Omega} [\phi_j(x) (W' * \phi_l(x)) + \phi_l(x) (W'* \phi_j(x))] \overline{\phi_k'(x)} \, dx \\
    &= \sum_{j=-L}^{L} \bar\mu_j \int_{\Omega} [\phi_j(x) \sqrt{d} w_l' \phi_l(x) + \phi_l(x) \sqrt{d} w_j' \phi_j(x) ] \overline{\phi_k'(x)} \, dx \\
    &= \frac{4\pi^2}{d\sqrt{d}} k \sum_{j=-L}^{L} \bar\mu_j ( l w_l + j w_j) \int_{\Omega} \phi_j(x) \phi_l(x) \phi_{-k}(x) \, dx \\
    &= \frac{4\pi^2}{d^2} k \sum_{j=-L}^{L} \bar\mu_j ( l w_l + j w_j) \int_{\Omega} \frac{1}{\sqrt{d}}e^{\frac{2\pi}{d}i(j+l-k)x} \, dx \\
    &= \frac{4\pi^2}{d^2} k \bar\mu_{k-l}(l w_l + (k-l)w_{k-l})\mathbbm{1}_{|k-l|\le L} ,
\end{split}
\]
where we used that
\[
W'(x) = \sum_{j=-\infty}^{\infty} w_j \phi_j'(x) = \sum_{j=-\infty}^{\infty} \frac{2\pi}{d} ij w_j \phi_j(x). 
\]

The next step is to calculate the matrix $T$.
We have
\[
\begin{split}
    T_{klm} &= \langle \phi_l W' * \phi_m, \phi_k' \rangle \\
    &= \frac{2\pi}{\sqrt{d}} im w_m \langle \phi_l \phi_m, \phi_k' \rangle \\
    &= \frac{4\pi^2}{d\sqrt{d}} k m w_m \int_{\Omega} \phi_l(x) \phi_m(x) \phi_{-k}(x) \, dx \\
    &= \frac{4\pi^2}{d^2} k m w_m \mathbbm{1}_{m=k-l} \mathbbm{1}_{|k-l| \le L}.     
\end{split}
\]
Given the sparsity of matrix $T$, we can compute $T : (a \otimes a)$ without storing $T$.

Notice that, for each $k$, 
\[
\sum_{l=-L}^{L} \sum_{m=-L}^{L} a_l(t) a_m(t) T_{klm} =  \frac{4\pi^2}{d^2} k \sum_{l=\max(k-L, -L)}^{\min(k+L, L)} a_l(t) a_{k-l}(t) (k-l) w_{k-l} =: T_k(a,a).
\]  

Then, we want to solve, for each $i$, and using the properties we discussed about this basis of functions,
\[
\begin{split}
    \dot{a}_i(t)  &= -\sum_{k=-L}^{L} a_k(t) (L_{ik} + \sigma D_{ik} + K_{ik}) - T_i(a,a).
\end{split}
\]

When including the control, notice that tensor $\Psi^{(j)}$ is similar to $L_G$ by replacing $G$ per $\alpha_j$ for each $j=1, \dots, m$. 
All the calculations remain the same.
For the operator $\mathcal{B}$, we need to apply $\mathcal{N}$ to $\bar\mu$.
When discretizing, we get 
\[
B_{kj} = -\int_{\Omega} \hat\mu_L(x) \alpha_j'(x) \overline{\phi_k'(x)} \, dx = -\sum_{l=-L}^L \bar\mu_l \int_{\Omega} \phi_l(x) \alpha_j'(x) \overline{\phi_k'(x)} \, dx \implies B_j = -\Psi^{(j)} \bar\mu.
\]

Another possibility is that $\alpha_j$ is not a periodic function, but $\nabla \alpha$ is a constant function.
In this case,
\[
\Psi_{kl}^{(j)} = \langle \phi_l, \phi_k' \rangle = \frac{2\pi}{d} ik\langle \phi_l, \phi_k \rangle = \frac{2\pi}{d} ik \mathbbm{1}_{k=l}.
\]

\subsubsection{2-D case}

In two dimensions, over the rectangle $[0,d_1] \times [0,d_2]$ with periodic boundary conditions, consider the orthonormal basis
\[
\phi_{k,l}(x,y) = \frac{1}{\sqrt{d_1 d_2}} e^{\frac{2\pi}{d_1}i k x} e^{\frac{2\pi}{d_2}i ly}.
\]

Consider a function $g$ written in this basis. 
So 
\[
g(x,y) = \sum_{k=-L}^L \sum_{l=-L}^L c_{k,l} \phi_{k,l}(x,y)
\]
over $[0,d_1] \times [0,d_2]$.
We know that
\[
c_{k,l} = \sqrt{\frac{1}{d_1 d_2}} \int_0^d f(x) e^{-\frac{2\pi}{d} k x i} \, dx \approx \frac{d}{N\sqrt{d}} \sum_{j=0}^{N-1} f(t_j) e^{-\frac{2\pi}{d} k t_j i}
\]
for points $0 \le t_0 < t_1 < \cdots < t_{N-1} \le d$. 
We consider the mid interval points $t_j = (j+1/2) d/N$, so
\[
\begin{split}
    c_k &\approx \frac{\sqrt{d}}{N} \sum_{j=0}^{N-1} f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)e^{-\frac{2\pi}{N} k (j+1/2) i} \\
    &= \frac{\sqrt{d}}{N} e^{-\frac{\pi}{N} k i}\sum_{j=0}^{N-1} f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)e^{-\frac{2\pi}{N} j k i} \\
    &=  \frac{\sqrt{d}}{N} e^{-\frac{\pi}{N} k i} FFT_k[(f_0, \dots, f_{N-1})],
\end{split}
\]
where $f_j = f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)$, for $k \ge 0$.
For $k < 0$, notice that $c_k = \overline{c_{-k}}$.

Let us calculate $W * \phi_k$ for a periodic function $W$.
Write 
\[
W(x) = \sum_{j=-\infty}^{\infty} w_j \phi_j (x).
\]
Then,
\[
W * \phi_k(x) = \sum_{j=-\infty}^{\infty} w_j \int_{\Omega} \phi_j(x-y) \phi_k(y) \, dy = \sum_{j=-\infty}^{\infty} w_j  \phi_j * \phi_k(x) = \sqrt{d} w_k \phi_k(x)
\]
where $w_k$ can be computed similarly using the FFT, as for $\mu_0$.

Using this fact to calculate $\bar\mu_i$,
\[
\bar\mu_i = \frac{1}{\hat{Z}} \int_{\Omega} \exp\left\{-\frac{G(x)}{\sigma} - \frac{\sqrt{d}}{\sigma}\sum_{k=-L}^L \bar\mu_k w_k \phi_k(x)\right\} \phi_i^*(x) \, dx.
\]

Let us find a better way to compute $L_G$.
Notice that \( \phi_k'(x) = \frac{2\pi}{d} ik \phi_k(x) \).
By definition,
\[
[L_G]_{kl} = \int_{\Omega} \phi_l(x) G'(x) \overline{\phi_k'(x)} \,dx = -\frac{2\pi}{d} i k \int_{\Omega} \phi_l(x) G'(x) \phi_{-k}(x) \,dx.
\]
Write 
\[
G(x) = \sum_{j=-\infty}^{\infty} g_j \phi_j(x) \implies G'(x) = \sum_{j=-\infty}^{\infty} g_j \phi_j'(x) = \sum_{j=-\infty}^{\infty} \frac{2\pi}{d} ij g_j \phi_j(x).
\]
Then, we get that 
\[
\begin{split}
    [L_G]_{kl} &= \frac{4\pi^2}{d^2} k \sum_{j=-\infty}^{\infty} j g_j \int_{\Omega} \phi_l(x) \phi_j(x) \phi_{-k}(x) \, dx \\
    &= \frac{4\pi^2}{d^2\sqrt{d}} k  \sum_{j=-\infty}^{\infty} j g_j \int_{\Omega} \frac{1}{\sqrt{d}} e^{\frac{2\pi}{d} i (l+j-k)x} \, dx \\
    &= \frac{4\pi^2}{d^2\sqrt{d}} k (k-l) g_{k-l}.
\end{split}
\]

We move to compute matrix $D$, which is given by
\[
D_{kl} = \int_{\Omega} \phi_l'(x) \overline{\phi_k'(x)} \, dx = \frac{4 \pi^2}{d^2} kl \int_{\Omega} \phi_{l}(x) \phi_{-k}(x) \, dx = \frac{4 \pi^2}{d^2} kl \mathbbm{1}_{k=l}.
\]

Let us do this for matrix $K$.
By definition, 
\[
\begin{split}
    K_{kl} &= \int_{\Omega} [\bar{\mu}_L(x) (W' * \phi_l(x)) + \phi_l(x) (W'*\bar\mu_L(x))] \overline{\phi_k'(x)} \, dx \\
    &= \sum_{j=-L}^{L} \bar\mu_j \int_{\Omega} [\phi_j(x) (W' * \phi_l(x)) + \phi_l(x) (W'* \phi_j(x))] \overline{\phi_k'(x)} \, dx \\
    &= \sum_{j=-L}^{L} \bar\mu_j \int_{\Omega} [\phi_j(x) \sqrt{d} w_l' \phi_l(x) + \phi_l(x) \sqrt{d} w_j' \phi_j(x) ] \overline{\phi_k'(x)} \, dx \\
    &= \frac{4\pi^2}{d\sqrt{d}} k \sum_{j=-L}^{L} \bar\mu_j ( l w_l + j w_j) \int_{\Omega} \phi_j(x) \phi_l(x) \phi_{-k}(x) \, dx \\
    &= \frac{4\pi^2}{d^2} k \sum_{j=-L}^{L} \bar\mu_j ( l w_l + j w_j) \int_{\Omega} \frac{1}{\sqrt{d}}e^{\frac{2\pi}{d}i(j+l-k)x} \, dx \\
    &= \frac{4\pi^2}{d^2} k \bar\mu_{k-l}(l w_l + (k-l)w_{k-l})\mathbbm{1}_{|k-l|\le L} ,
\end{split}
\]
where we used that
\[
W'(x) = \sum_{j=-\infty}^{\infty} w_j \phi_j'(x) = \sum_{j=-\infty}^{\infty} \frac{2\pi}{d} ij w_j \phi_j(x). 
\]

The next step is to calculate the matrix $T$.
We have
\[
\begin{split}
    T_{klm} &= \langle \phi_l W' * \phi_m, \phi_k' \rangle \\
    &= \frac{2\pi}{\sqrt{d}} im w_m \langle \phi_l \phi_m, \phi_k' \rangle \\
    &= \frac{4\pi^2}{d\sqrt{d}} k m w_m \int_{\Omega} \phi_l(x) \phi_m(x) \phi_{-k}(x) \, dx \\
    &= \frac{4\pi^2}{d^2} k m w_m \mathbbm{1}_{m=k-l} \mathbbm{1}_{|k-l| \le L}.     
\end{split}
\]
Given the sparsity of matrix $T$, we can compute $T : (a \otimes a)$ without storing $T$.

Notice that, for each $k$, 
\[
\sum_{l=-L}^{L} \sum_{m=-L}^{L} a_l(t) a_m(t) T_{klm} =  \frac{4\pi^2}{d^2} k \sum_{l=\max(k-L, -L)}^{\min(k+L, L)} a_l(t) a_{k-l}(t) (k-l) w_{k-l} =: T_k(a,a).
\]  

Then, we want to solve, for each $i$, and using the properties we discussed about this basis of functions,
\[
\begin{split}
    \dot{a}_i(t)  &= -\sum_{k=-L}^{L} a_k(t) (L_{ik} + \sigma D_{ik} + K_{ik}) - T_i(a,a).
\end{split}
\]

When including the control, notice that tensor $\Psi^{(j)}$ is similar to $L_G$ by replacing $G$ per $\alpha_j$ for each $j=1, \dots, m$. 
All the calculations remain the same.
For the operator $\mathcal{B}$, we need to apply $\mathcal{N}$ to $\bar\mu$.
When discretizing, we get 
\[
B_{kj} = -\int_{\Omega} \hat\mu_L(x) \alpha_j'(x) \overline{\phi_k'(x)} \, dx = -\sum_{l=-L}^L \bar\mu_l \int_{\Omega} \phi_l(x) \alpha_j'(x) \overline{\phi_k'(x)} \, dx \implies B_j = -\Psi^{(j)} \bar\mu.
\]

Another possibility is that $\alpha_j$ is not a periodic function, but $\nabla \alpha$ is a constant function.
In this case,
\[
\Psi_{kl}^{(j)} = \langle \phi_l, \phi_k' \rangle = \frac{2\pi}{d} ik\langle \phi_l, \phi_k \rangle = \frac{2\pi}{d} ik \mathbbm{1}_{k=l}.
\]

\subsubsection{Ideas to implement}

\begin{enumerate}
    \item I can reduce the system size by using symmetry in the complex numbers, such as, \(a_{-k} = \overline{a_k}\). 
    For this to work, I must redo all matrices considering the system reduction: 
    For the linear case, it is not hard. 
    Suppose $x = (x^{0}, x^{1}, x^{2})$ and $x' = Ax$ for a matrix $A$. 
    Consider that $A$ can be divided in six parts and $x^{2} = \overline{x^{1}}$. 
    We suppose that $x_0$ is real.
    Then,
    \begin{align*}
        \dot{\Re(x^{1})} &= \Re(A_0) x_0 + \Re(A_1) \Re(x^1) - \Im(A_1) \Im(x^1) + \Re(A_2) \Re(x^1) + \Im(A_2) \Im(x^1) \\
        \dot{\Im(x^{1})}&= \Im(A_0) x_0 + \Re(A_1) \Im(x^1) + \Im(A_1) \Re(x^1) - \Re(A_2) \Im(x^1) - \Im(A_2) \Re(x^1),
    \end{align*}
    which is an affine system if we suppose that $x_0$ is known.
    For the quadratic transformation $T_k$, we need more work because 
    \[
    \begin{split}
        T_k(a,a) &= -\sum_{s=1}^{L-k} (\Re(a_{k+s}) + i \Im(a_{k+s}))(\Re(a_{s}) - i \Im(a_{s}))s(\Re(w_{s}) - i \Im(w_{s})) \\ &+ \sum_{s=1}^{k-1} (\Re(a_{k-s}) + i \Im(a_{k-s}))(\Re(a_{s}) + i \Im(a_{s}))s(\Re(w_{s}) + i \Im(w_{s})) \\ &+ \sum_{s=k}^{L} (\Re(a_{s-k}) - i \Im(a_{s-k}))(\Re(a_{s}) + i \Im(a_{s}))s(\Re(w_{s}) + i \Im(w_{s})).
    \end{split}
    \]


    \item matrix $L_G$, $D$ and $K$ have special structures.
    In particular, $L_G a$ can be viewed as a convolution of $a$ instead of a matrix product. 
    Therefore, we can apply the Fast Fourier Transform and reduce the cost from $O(L^2)$ to $O(L \log L)$.
\end{enumerate}


\newpage

\bibliographystyle{plainnat}
\bibliography{biblio}

\end{document}

\subsubsection{1-D case}

In one dimension, over the interval $[0,d]$ with periodic boundary conditions, consider the orthonormal basis
\[
\phi_k(x) = \begin{cases}
    \sqrt{\frac{2}{d}} \cos\left(\frac{2k\pi x}{d} \right), &k \ge 1, \\
    \sqrt{\frac{1}{d}}, &k =0, \\
    \sqrt{\frac{2}{d}} \sin\left(\frac{2|k|\pi x}{d} \right), &k \le -1.
\end{cases}
\]

An easy way to project a function on the Fourier basis uses the so-called Fast Fourier Transform.
If over an interval $[0,d]$,
\[
g(x) = \sum_{k=-L}^L c_k \psi_k(x), \quad \psi_k(x) = \sqrt{\frac{1}{d}}e^{\frac{2\pi}{d} k x i},
\]
then we know that
\[
c_k = \sqrt{\frac{1}{d}} \int_0^d f(x) e^{-\frac{2\pi}{d} k x i} \, dx \approx \frac{d}{N\sqrt{d}} \sum_{j=0}^{N-1} f(t_j) e^{-\frac{2\pi}{d} k t_j i}
\]
for points $0 \le t_0 < t_1 < \cdots < t_{N-1} \le d$. 
We consider the mid interval points $t_j = (j+1/2) d/N$, so
\[
\begin{split}
    c_k &\approx \frac{\sqrt{d}}{N} \sum_{j=0}^{N-1} f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)e^{-\frac{2\pi}{N} k (j+1/2) i} \\
    &= \frac{\sqrt{d}}{N} e^{-\frac{\pi}{N} k i}\sum_{j=0}^{N-1} f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)e^{-\frac{2\pi}{N} j k i} \\
    &=  \frac{\sqrt{d}}{N} e^{-\frac{\pi}{N} k i} FFT_k[(f_0, \dots, f_{N-1})],
\end{split}
\]
where $f_j = f\left(\left(j+\frac{1}{2}\right)\frac{d}{N}\right)$.
After computing $c_k$ values, by Euler relation, we obtain the values for the writing using cosine and sine functions.
That is, if
\[
g(x) = \sum_{k=-L}^L a_k \phi_k(x),
\]
we have $a_0 = c_0$ and $a_n = 2\operatorname{Re}(c_n)/\sqrt{2}$ for $n \ge 1$ and $a_n = 2\operatorname{Im}(c_n)/\sqrt{2}$ for $n \le -1$.

Let us calculate $W * \phi_k$ for a periodic function $W$.
Write 
\[
W(x) = \sum_{j=-\infty}^{\infty} w_j \phi_j (x).
\]
Then,
\[
\begin{split}
    W * \phi_k(x) &= \sum_{j=-\infty}^{\infty} w_j \int_{\Omega} \phi_j(x-y) \phi_k(y) \, dy \\ 
    &= \sum_{j=-\infty}^{\infty} w_j  \phi_j * \phi_k(x) = \begin{cases}
    \sqrt{\frac{d}{2}}(\operatorname{sgn}(k) w_k \phi_{|k|}(x) + w_{-k}\phi_{-|k|}(x)), &k\neq 0 \\
    w_0 \sqrt{d}\phi_0(x), &k=0 
    \end{cases}
\end{split}
\]
where $w_k$ can be computed similarly using the FFT, as for $\mu_0$ and we can pre-compute $\phi_j * \phi_k$ as:
\[
\phi_j * \phi_k = \sqrt{\frac{d}{2}} \left( \operatorname{sgn}(j)\phi_{|j|} \mathbbm{1}_{j=k} + \phi_{-|j|} \mathbbm{1}_{j=-k} \right),
\]
for $j \neq 0$. 
If $j=0$ and $k \neq 0$, this composition is $0$, but if $j=0$ and $k=0$, it is $1$.

Using this fact to calculate $\bar\mu_i$,
\[
\bar\mu_i = \frac{1}{\hat{Z}} \int_{\Omega} \exp\left\{-G(x) -  \left(\sum_{k=-L, k\neq 0}^L \bar\mu_k \sqrt{\frac{d}{2}}\left(\operatorname{sgn}(k) w_k \phi_{|k|}(x) + w_{-k}\phi_{-|k|}(x)\right) + \bar\mu_0 w_0\right)\right\} \phi_i^*(x) \, dx,
\]

Let us find a better way to compute $L_G$.
\[
\phi_k'(x) = \begin{cases}
    -\frac{2k\pi}{d} \sqrt{\frac{2}{d}} \sin\left(\frac{2k\pi x}{d} \right) = -\frac{2k\pi}{d} \phi_{-k}(x), &k \ge 1, \\
    0, &k =0, \\
    \frac{2|k|\pi}{d}\sqrt{\frac{2}{d}} \cos\left(\frac{2|k|\pi x}{d} \right) = \frac{2|k|\pi}{d} \phi_{-k}(x), &k \le -1.
\end{cases}
\]

By definition,
\[
[L_G]_{ij} = \int_{\Omega} \phi_j(x) G'(x) \phi_i'(x) \,dx = -2\frac{i\pi}{d} \int_{\Omega} \phi_j(x) G'(x) \phi_{-i}(x) \,dx,
\]
which implies that we can first project $\phi_j(x) G'(x)$ onto the Fourier basis, reverse the coefficients and then multiply by $-2i\pi/d$.
But we can go further.
Write 
\[
G'(x) = \sum_{k=-\infty}^{\infty} g_k \phi_k(x).
\]
Then, we get that 
\[
[L_G]_{ij} = -\frac{2   i\pi}{d}\sum_{k=-\infty}^{\infty} g_k \int_{\Omega} \phi_j(x) \phi_k(x) \phi_{-i}(x) \, dx, 
\]
which can be separated into six cases depending on the sign of $i$ and $j$.

\begin{enumerate}
    \item If $i=0$, then this integral will be always zero when multiplied by $i$.

    \item If $i\neq 0$, but $j = 0$, then
    \[
    \int_{\Omega} \phi_j(x) \phi_k(x) \phi_{-i}(x) \, dx = \sqrt{\frac{1}{d}} \int_{\Omega} \phi_k(x) \phi_{-i}(x) \, dx = \sqrt{\frac{1}{d}}\mathbbm{1}_{\{k=-i\}}.
    \]

    \item If $i < 0$ and $j > 0$, then 
    \[
    \begin{split}
             \int_{\Omega} \phi_j(x) \phi_k(x) \phi_{-i}(x) \, dx &= \frac{2}{d} \int_{\Omega} \cos\left(\frac{j\pi x}{d}\right)\cos\left(\frac{-i\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \int_{\Omega} \cos\left(\frac{(j+i)\pi x}{d}\right) \phi_{k}(x) \, dx + \frac{1}{d} \int_{\Omega} \cos\left(\frac{(j-i)\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{|j+i|=k} + \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{|j-i|=k}.
    \end{split}
    \]

    \item If $i < 0$ and $j < 0$, then 
    \[
    \begin{split}
             \int_{\Omega} \phi_j(x) \phi_k(x) \phi_{-i}(x) \, dx &= \frac{2}{d} \int_{\Omega} \sin\left(\frac{|j|\pi x}{d}\right)\cos\left(\frac{-i\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \int_{\Omega} \sin\left(\frac{(-i+|j|)\pi x}{d}\right) \phi_{k}(x) \, dx - \frac{1}{d} \int_{\Omega} \sin\left(\frac{(-i-|j|)\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \operatorname{sgn}(-i + |j|) \mathbbm{1}_{|-i+|j||=|k|} - \frac{1}{d} \sqrt{\frac{d}{2}} \operatorname{sgn}(-i - |j|) \mathbbm{1}_{|-i-|j||=|k|} \\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{|i+j|=-k} - \frac{1}{d} \sqrt{\frac{d}{2}} \operatorname{sgn}(-i+j)\mathbbm{1}_{|j-i|=-k}
    \end{split}
    \]

    \item If $i > 0$ and $j > 0$, then 
    \[
    \begin{split}
             \int_{\Omega} \phi_j(x) \phi_k(x) \phi_{-i}(x) \, dx &= \frac{2}{d} \int_{\Omega} \cos\left(\frac{j\pi x}{d}\right)\sin\left(\frac{i\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \int_{\Omega} \sin\left(\frac{(j+i)\pi x}{d}\right) \phi_{k}(x) \, dx - \frac{1}{d} \int_{\Omega} \sin\left(\frac{(j-i)\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \operatorname{sgn}(j + i) \mathbbm{1}_{|j+i|=|k|} - \frac{1}{d} \sqrt{\frac{d}{2}} \operatorname{sgn}(j - i) \mathbbm{1}_{|j-i|=|k|}\\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{|j+i|=-k} - \frac{1}{d} \sqrt{\frac{d}{2}} \operatorname{sgn}(j -i) \mathbbm{1}_{|j-i|=-k}
    \end{split}
    \]

    \item If $i > 0$ and $j < 0$, then
    \[
    \begin{split}
             \int_{\Omega} \phi_j(x) \phi_k(x) \phi_{-i}(x) \, dx &= \frac{2}{d} \int_{\Omega} \sin\left(\frac{|j|\pi x}{d}\right)\sin\left(\frac{i\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \int_{\Omega} \cos\left(\frac{(|j|-i)\pi x}{d}\right) \phi_{k}(x) \, dx - \frac{1}{d} \int_{\Omega} \cos\left(\frac{(|j|+i)\pi x}{d}\right) \phi_{k}(x) \, dx \\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{||j|-i|=k} - \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{||j|+i|=k} \\
             &= \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{|j+i|=k} - \frac{1}{d} \sqrt{\frac{d}{2}} \mathbbm{1}_{|j-i|=k}.
    \end{split}
    \]

\end{enumerate}

We move to compute matrix $D$, which is given by
\[
D_{ij} = \int_{\Omega} \phi_j'(x) \phi_i'(x) \, dx = 4\frac{ij \pi^2}{d^2} \int_{\Omega} \phi_{-j}(x) \phi_{-i}(x) \, dx = 4\frac{ij \pi^2}{d^2} \mathbbm{1}_{i=j}
\]

Let us do matrix $K$.
By definition, 
\[
\begin{split}
    K_{ij} &= \int_{\Omega} [\bar{\mu}_L(x) (W' * \phi_j(x)) + \phi_j(x) (W'*\bar\mu_L(x))] \phi_i'(x) \, dx \\
    &= \sum_{k=-L}^{L} \bar\mu_k \int_{\Omega} [\phi_k(x) (W' * \phi_j(x)) + \phi_j(x) (W'* \phi_k(x))] \phi_i'(x) \, dx \\
    &= \sum_{k=-L}^{-1} \bar\mu_k \int_{\Omega} [\phi_k(x) \sqrt{\frac{d}{2}}\left( \operatorname{sgn}(j) w_j \phi_{|j|}(x) + w_{-j} \phi_{-|j|} \right)  + \phi_j(x)  \\
    &= \sum_{k=-L}^{L} \bar\mu_k \int_{\Omega} [\phi_k(x) w_j' \phi_j(x) + \phi_j(x) w_k' \phi_k(x)] \phi_i'(x) \, dx \\
    &= \sum_{k=-L}^L \bar\mu_k \left(  w_j' \int_{\Omega} \phi_k(x) \phi_j(x) \phi_i'(x) \, dx  + w_k' \int_{\Omega} \phi_k(x) \phi_j(x) \phi_i'(x) \, dx \right) \\
    &= \sum_{k=-L}^L \bar\mu_k \left(-\frac{2j\pi}{d} w_{-j} \frac{2i\pi}{d}\int_{\Omega} \phi_k(x) \phi_j(x) \phi_{-i}(x) \, dx - \frac{2k\pi}{d} w_{-k} \frac{2i\pi}{d}\int_{\Omega} \phi_k(x) \phi_j(x) \phi_{-i}(x) \, dx \right) \\
    &= -\frac{4i\pi^2}{d^2}\sum_{k=-L}^L \bar\mu_k \int_{\Omega} \phi_k(x) \phi_j(x) \phi_{-i}(x) \, dx \left( j w_{-j} + k w_{-k}  \right),
\end{split}
\]
where we used the previous known fact that $W' * \phi_k(x) = w_k' \phi_k(x)$ with $w_k' = 2 k \pi w_{-k} / d$.
As an observation, 
\[
W'(x) = \frac{d}{dx} \sum_{j=-\infty}^{\infty} w_j \phi_j(x) = -\frac{2\pi}{d} \sum_{j=-\infty}^{\infty} w_j j \phi_{-j}(x) = \frac{2\pi}{d} \sum_{l=-\infty}^{\infty} w_{-l} l \phi_{l}(x).
\]
Notice that we already have all the necessary estimates to compute $K$.